### 2022-09-28 Thesis/text/diskussion.tex
fertig?!

This thesis demonstrated how to incorporate advanced behavior planning in ROS2 mobile robots. The goal was to achieve higher levels of autonomy and robustness by reducing the number of scenarios in which a human operator is needed. Implementing the system supervisor component improves the system's robustness and provides a safe basis for executing more intelligent behaviors. The implemented behaviors to react to sensor failures and robot collisions are valuable improvements to the system and improve current ROS2 mobile robots significantly. The layout of the behavior tree is structured to allow more complex behaviors to be safely executed because possible scenarios that interfere with a correct environment representation are checked and dealt with before complex behaviors get executed based on that representation. 
The implemented layout choice for the tree is not mentioned in the literature. However, it has proven to be a sensible approach to model robot behavior safely and robustly. 
%The implemented behaviors and the system supervision component have improved the behavior in the simulated scenarios. The system can react and recover from multiple scenarios with additional components.

Furthermore, the system architecture is a good base for further development without facing problems or limitations in the later stages of developing behaviors. The architecture shows a novel way that works with the asynchronous ROS2 executor properties and allows fast execution of both the behavior planner and the system components.

Nevertheless, the behavior tree design could be modified to use blackboard values more effectively to increase the information flow between action nodes in the tree. Also, the tree does not utilize the reactive control flow nodes, and no action node in the tree is designed to allow multiple threads at the current state. Incorporating so-called stateful nodes into the tree's structure to allow co-routines would increase the possibility of creating better and more deliberative behaviors.

@@

@@

Lastly, the results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

>>>>

This thesis demonstrated how to incorporate advanced behavior planning in ROS2 mobile robots. The goal was to achieve higher levels of autonomy and robustness by reducing the number of scenarios in which a human operator is needed. Implementing the system supervisor component improved the system's robustness and provided a safe basis for executing more intelligent behaviors. The implemented behaviors to react to sensor failures and robot collisions were valuable improvements to the system and improved current ROS2 mobile robots significantly. The layout of the behavior tree was structured to allow more complex behaviors to be safely executed because possible scenarios that interfere with a correct environment representation were checked and dealt with before complex behaviors get executed based on that representation. 
The implemented layout choice for the tree was not mentioned in previous literature. However, it has proven to be a sensible approach to model robot behavior safely and robustly. 
%The implemented behaviors and the system supervision component have improved the behavior in the simulated scenarios. The system can react and recover from multiple scenarios with additional components.

Furthermore, the system architecture is a good base for further development without facing problems or limitations in the later stages of developing behaviors. The architecture showed a novel way that works with the ROS2 executor properties and allowed fast execution of both the behavior planner and the system components.

Nevertheless, the behavior tree design could have been modified to use blackboard values more effectively to increase the information flow between action nodes in the tree. Also, the tree did not utilize the reactive control flow nodes, and no action node in the tree was designed to allow multiple threads at the current state. Incorporating so-called stateful nodes into the tree's structure to allow co-routines would have increased the possibility of creating better and more deliberative behaviors.

@@

The results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

@@

### 2022-09-28 Thesis/text/ergebnisse.tex
modification to line spacing, changed bt pictures
fertig?!

The robot is put into specifically designed scenarios that should trigger behavior responses from the behavior tree component. The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.

The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}. The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.

@@

The robot location and navigation goals are defined in another ROS Node, which programmatically spawns the robot into the environment and sets the goal this way. This way, the same scenario can be run with and without behavior planning. To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in section \ref{subsec:gazebo_driver}, was shut down to emulate the sensor's failure.

@@

\begin{table}[ht]
	\centering
	\caption{Sensor Scenarios}
	\label{tab:sensor_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
			\hline
			\textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
			\hline
			Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
			\hline
			Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\hline
			Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
			IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
			\hline
			IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\hline
			IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
			Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
			\hline
			Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\hline
			Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
		\end{tabular}
	}
\end{table}

@@

\begin{itemize}
	\item For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.

	\item Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered. The BT has to recover from the inability to plan in this scenario.

	\item NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal. This scenario tests the ability to find alternative goals and navigate to a nearby goal.

	\item In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered. The forced collision tests the ability to escape a collision and resume navigation afterward. A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.

	\item Collision$\_$2 scenario is testing the same behavior but during active navigation. A flat obstacle is spawned in the path of the robot. The obstacle is too low to be detected by the laserscanner, and a collision occurs. The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.

	\item The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge. The simulated battery is emptied to a low level, and a new navigation goal is set. If the battery charge runs below zero percent during the navigation, the test fails. 
\end{itemize}

\begin{table}[ht]
	\centering
	\caption{Behavior Scenarios}
	\label{tab:behavior_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
		\hline
		\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
	 	\hline	
	 	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still & 
	Reset system, calculate path to goal, reach goal \\
		\hline
		NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
		\hline
		Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
		\hline
		Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) & 
	Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
		\hline
		Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
		\hline
		Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
		\hline	
		\end{tabular}
	}
\end{table}

>>>>

The robot is put into specifically designed scenarios to trigger behavior responses from the behavior tree component. The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.

The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}. The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided by the Nav2 map server.

@@

The robot location and navigation goals are defined in another ROS node, which programmatically spawns the robot into the environment and sets the goal this way. This way, the same scenario can be run with and without behavior planning. To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in section \ref{subsec:gazebo_driver}, was shut down to emulate the sensor's failure.

@@

\begin{table}[ht]
	\centering
	\caption{Sensor Scenarios}
	\label{tab:sensor_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.99\textwidth}{!}{%
		\begin{tabular}{| m{0.11\textwidth}| m{0.1\textwidth} | m{0.183\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
			\hline
			\textbf{Sensor Tested} & \textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
			\hline
			 & Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
			\cline{2-5}
			Lidar &Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\cline{2-5}
			& Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
			 & IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
			\cline{2-5}
			IMU &IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\cline{2-5}
			&IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
			 & Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
			\cline{2-5}
			Odometry &Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
			\cline{2-5}
			&Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
			\hline
		\end{tabular}
	}
\end{table}

\begin{table}[ht]
	\centering
	\caption{Behavior Scenarios}
	\label{tab:behavior_scenarios}
	\renewcommand{\arraystretch}{1.5}
	\resizebox{0.9\textwidth}{!}{%
		\begin{tabular}{| m{0.1\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
			\hline
			\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
			\hline	
			NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still & 
			Reset system, calculate path to goal, reach goal \\
			\hline
			NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
			\hline
			Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
			\hline
			Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) & 
			Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
			\hline
			Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
			\hline	
		\end{tabular}
	}
\end{table}

@@

For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.

Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered. The BT has to recover from the inability to plan in this scenario.

NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal. This scenario tests the ability to find alternative goals and navigate to a nearby goal.

In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered. The forced collision tests the ability to escape a collision and resume navigation afterward. A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.

Collision$\_$2 scenario is testing the same behavior but during active navigation. A flat obstacle is spawned in the path of the robot. The obstacle is too low to be detected by the laserscanner, and a collision occurs. The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.

The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge. The simulated battery is emptied to a low level, and a new navigation goal is set. If the battery charge runs below zero percent during the navigation, the test fails.
